Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-13 main.tex
Addressed the second reviewer's comments on the two fall modes.

\subsection{Protocol}
%
We recruited 26 able-bodied young adults between 20 and 36 years old (mean 26
years~\footnote{Four participants' ages were lost but did fall within the
stated range.}) to participate in the experiments. The participants were all
confident in their cycling skills and had cycled at least once in the last
month. All participants consented to the experiment and could decline to
continue at any time. The study was approved by Delft University of
Technology's Human Research Ethics Council (record \#3897). The participants
were divided into two groups.  The first group of fifteen participants
performed the protocol at a belt speed of 10~\si{\kph}(2.8~\si{\mps}) with the
gain factor set to \(\kappa=3.9\) and two weeks later the second group of
eleven participants performed the protocol at a belt speed of 6~\si{\kph}
(1.7~\si{\mps}) with the gain factor set to \(\kappa=5.2\).

Participants wore a helmet and they were attached to the ceiling via a fall
safety harness, Figure~\ref{fig:participant-in-set-up}. The harness allowed
natural free movement pre-fall. The participants practiced falling in the
harness at least once and then practiced riding on the treadmill until they
indicated they were comfortable enough to have perturbations applied. For most,
this was less than a 10~\si{\minute} warm up. We then asked the rider to ride
for 90~\si{\second}, while attempting to maintain the location of their front
wheel on the center line of the treadmill as a baseline measure before the
perturbations. We then applied perturbations in random directions (clockwise or
counter-clockwise), starting at 20~\si{\newton} and increasing the magnitude by
30~\si{\newton} until the participants fell. We defined a ``fall'' on the
treadmill by two criteria: 1) the rider removes their foot from the pedal and
places it on the ground or 2) the bicycle wheel exceeds the width of the
treadmill belt. These were the two actions which necessitated stopping the
treadmill belt. We informed the participants beforehand that we would stop the
treadmill in these situations. Placing a foot down on the fall side of the
bicycle is a natural reaction when steering or other body movements are not
sufficient to keep the bicycle upright, as shown in \citep{Gildea2024a}.
Exiting the width of the treadmill is akin to exiting the width of a curb-less
bicycle path and arises when the rider's control actions are not sufficient to
correct the heading of the vehicle in sufficient time.
Figure~\ref{fig:perturbation-sequence} shows an example resulting motion from a
perturbation. We logged the force magnitude that caused the first fall to
characterize that participant's \emph{perturbation resistance threshold}.
%
\begin{figure}
  \centering
  \subcaptionbox{Before perturbation}{\includegraphics[width=60mm]{figures/start.png}}
  \subcaptionbox{During perturbation}{\includegraphics[width=60mm]{figures/during.png}}
  \subcaptionbox{After perturbation}{\includegraphics[width=60mm]{figures/after.png}}
  \subcaptionbox{Recovery from perturbation}{\includegraphics[width=60mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}

>>>>

\subsection{Protocol}
%
We recruited 26 able-bodied young adults between 20 and 36 years old (mean 26
years~\footnote{Excluding four participants' ages that were misplaced but that
did fall within the stated range.}) to participate in the experiments. The
participants were all confident in their cycling skills and had cycled at least
once in the last month. All participants consented to the experiment and could
decline to continue at any time. The study was approved by Delft University of
Technology's Human Research Ethics Council (record \#3897). The participants
were divided into two groups. The first group of fifteen participants
performed the protocol at a belt speed of 10~\si{\kph}(2.8~\si{\mps}) with the
gain factor set to \(\kappa=3.9\) and two weeks later the second group of
eleven participants performed the protocol at a belt speed of 6~\si{\kph}
(1.7~\si{\mps}) with the gain factor set to \(\kappa=5.2\).

Participants wore a helmet and they were attached to the ceiling via a fall
safety harness, Figure~\ref{fig:participant-in-set-up}. The harness allowed
natural free movement pre-fall. The participants practiced falling in the
harness at least once and then practiced riding on the treadmill until they
indicated they were comfortable enough to have perturbations applied. For most,
this was less than a 10~\si{\minute} warm up. We then asked the rider to ride
for 90~\si{\second}, while attempting to maintain the location of their front
wheel on the center line of the treadmill as a baseline measure before the
perturbations. We then applied perturbations in random directions (clockwise or
counter-clockwise), starting at 20~\si{\newton} and increasing the magnitude by
30~\si{\newton} until the participants fell. We defined a ``fall'' on the
treadmill by two criteria: 1) the rider removes their foot from the pedal and
places it on the ground or 2) the bicycle wheel exceeds the width of the
treadmill belt.
These two fall criteria are not strictly the same and the latter is more likely
to happen at higher speeds.
We informed the participants beforehand that we would stop the treadmill in
these situations.
These were the two actions which necessitated stopping the treadmill belt and
the simulate real fall modes.
Placing a foot down on the fall side of the bicycle is a natural reaction when
steering or other body movements are not sufficient to keep the bicycle
upright, as shown in \citep{Gildea2024a}.
Exiting the width of the treadmill is akin to exiting the width of a curb-less
bicycle path and arises when the rider's control actions are not sufficient to
correct the heading of the vehicle in sufficient time.
Figure~\ref{fig:perturbation-sequence} shows an example resulting motion from a
perturbation. We logged the force magnitude that caused the first fall to
characterize that participant's \emph{perturbation resistance threshold}.
%
\begin{figure}
  \centering
  \subcaptionbox{Before perturbation}{\includegraphics[width=60mm]{figures/start.png}}
  \subcaptionbox{During perturbation}{\includegraphics[width=60mm]{figures/during.png}}
  \subcaptionbox{After perturbation}{\includegraphics[width=60mm]{figures/after.png}}
  \subcaptionbox{Recovery from perturbation}{\includegraphics[width=60mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}
